### 2015-05-12 tex/sharp-sigma-one-plus.tex
22 abr
29 abr 1
12 may

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm} 
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}

@@

\def\E1{\#\Sigma_1^{+}}

@@

@@

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.

Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables. A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\

The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.

@@

@@

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm} 
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{subfiles}
\usepackage{bm}

@@

\def\E1{\#\Sigma_1[\mbox{{\sc FO}}]}

@@

\def\np{\mbox{{\sc NP}}}

@@

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.

A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\

The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \Truc \mid f(\A) > 0\}$.

@@

For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
\begin{eqnarray*}
(f \dotminus g)(\A) =
\begin{cases}
f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
0, & \text{if }f(\A) \leq g(\A).
\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.

@@

\bibliographystyle{plain}
\bibliography{SalujaST95}

### 2015-05-12 tex/th1.tex
12 may

\begin{theo} \label{decisionptime}
The decision version of a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$. Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$. Note that $m$ and $k$ are fixed values, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$. Let $\varphi_\A(\X)$ be defined as follows:
\begin{eqnarray*}
\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
\end{eqnarray*}
Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$. Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction. We can do this since checking $\A\models\varphi(\d_i,\X,\e_i)$ can be done in polynomial time for each $i\in\{1,\ldots,n\}$. Note as well that $\varphi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$. Then, let $f'$ be defined as follows:
\begin{eqnarray*}
f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
\end{eqnarray*}
Clearly $f' \in\#\Sigma_0$. Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$. However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time. Therefore, the decision version of $f$ is in {\sc P}.
\end{proof}

>>>>

%!TEX root = sharp-sigma-one-plus.tex
\begin{theo} \label{decisionptime}
The decision problem associated to a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$. Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$. Note that the values of $m$ and $k$ come from $\varphi(\x,\X,\z)$, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$. Slightly abusing notation, let $\varphi_\A(\X)$ be defined as follows:
\begin{eqnarray*}
\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
\end{eqnarray*}
Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ if and only if $f(\A)>0$. Let $\psi_\A(\X)$ be the following formula: For each FO-formula $\beta(\x,\z)$ in $\varphi(\x,\X,\z)$ we check if $\A \models \beta(\d_i,\e_j)$, for every $i\in\{1,\ldots,\ell\}$ and $j\in\{1,\ldots,m\}$. If it is true, then we replace $\beta(\d_i,\e_j)$ with a tautology, and otherwise, with a contradiction. Let us give an example on how this would be done: \\

Let $\varphi(\x,\X,\z) = \exists x_1\exists x_2( X_1(x_2) \wedge ( X_2(z_1,x_1) \wedge \forall y( S_1(y,z_1) )  ))$. In this case, $\x = (x_1,x_2)$, $\X = (X_1,X_2)$ and $\z = (z_1)$. Let $\A = \langle A,S^\A,\leq^\A \rangle$, where $A = \{1,2\}$, $S^\A = \{(1,1),(1,2),(2,2)\}$, and $1 \leq^\A 2$. We generate all the possible assignments for $\x$ and $\z$. And our new formula is:
\begin{multline*}
\varphi_\A(\X) = (X_1(1) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  )\; \vee \\
(X_1(2) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  )\; \vee \\
(X_1(1) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  )\; \vee \\
(X_1(2) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  ).
\end{multline*}
Now we replace the satisfied subformulas:
\begin{multline*}
\psi_\A(\X) = (X_1(1) \wedge X_2(1,1) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,1) \wedge \top  ) \;\vee \\
(X_1(2) \wedge  X_2(1,1) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,1) \wedge \top  ) \;\vee \\
(X_1(1) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,2) \wedge \top  ) \;\vee \\
(X_1(2) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,2) \wedge \top  ),
\end{multline*}
which can be simplified even further:
\begin{multline*}
\psi_\A(\X) = (X_1(1) \wedge X_2(2,1)  ) \vee (X_1(2) \wedge  X_2(2,1) ) \;\vee \\ 
(X_1(1) \wedge  X_2(2,2)  ) \vee (X_1(2) \wedge  X_2(2,2)  )
\end{multline*}
Note that we can do this since checking $\A\models\varphi(\d_i,\X,\e_j)$ can be done in polynomial time for each $i,j$, as $\varphi$ is assumed to be fixed. Note as well that $\psi_\A(\X)$ is quantifier-free and its size is also polynomial to the size of $\A$. However, this still is not a valid $\L-$formula since there are no constants in $\L$. Let $P_a$ be a unary predicate for every $a\in A$. Now let $\zeta_\A(\X,\u)$ be obtained from $\psi_\A(\X)$ by adding the conjunct $P_{a_1}(u_{a_1}) \wedge \cdots \wedge P_{a_{\vert A \vert}}(u_{a_{\vert A \vert}})$, where $u_a$ is a first order variable, and by replacing every constant $a\in A$ by $u_a$. In the example:
\begin{multline*}
\zeta_\A(\X,\u) = \big((X_1(u_1) \wedge X_2(u_2,u_1)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_1) )\; \vee \\ 
(X_1(u_1) \wedge  X_2(u_2,u_2)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_2)  )\big) \wedge P_1(u_1) \wedge P_2(u_2),
\end{multline*}
where $\u = (u_1,u_2)$. Now, let $\L^\prime = \L \cup \{P_a\mid a\in A\}$ and $\A^\prime = \langle A,S^\A, \P^{\A^\prime}_{a_1},\ldots,\P^{\A^\prime}_{a_{\vert A \vert}}, \leq^\A \rangle$, where $\P^{\A^\prime}_a = \{a\}$ for each $a\in A$. Moreover, let $f^\prime$ be defined as follows:
\begin{eqnarray*}
f^\prime(\A^\prime) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \zeta_\A(\P,\e) \} \mid.
\end{eqnarray*}
Clearly $f^\prime \in\#\Sigma_0$. Also, note that for every $\A$, $f(\A) > 0$ if and only if $f^\prime(\A^\prime) > 0$, so computing $f^\prime$ with the instance $\A^\prime$ is enough to solve the decision version of $f$. Therefore, given that any counting problem in $\#\Sigma_0$ is computable in polynomial time\cite{DBLP:journals/jcss/SalujaST95}, we conclude that the decision version of $f$ is in {\sc P}.
\end{proof}

### 2015-05-12 tex/th2.tex
12 may

\begin{theo}
If $\#\Sigma_1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theo}
\begin{proof}
Let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment (satisfying or not) to a 3DNF formula. Suppose that $f_{2^n}-f_{\#3DNF} \in \#\Sigma_1$. This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete. However, as we showed previously in Theorem \ref{decisionptime}, it is also in P. Then, co-NP $\subseteq$ P.
\end{proof}

>>>>

\begin{theo}
If $\#\Sigma_1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theo}
\begin{proof}
Suppose that $\#\Sigma_1$ is closed under substraction, that is, for each pair of functions $f,g\in\#\Sigma_1,$ there is an $h\in\#\Sigma_1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
\begin{eqnarray*}
S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
\end{eqnarray*}
Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (X(x) \vee \neg X(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. As we supposed initially, let $h\in\#\Sigma_1$ such that $f_{all}-f_{\#3DNF} = h$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete. However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$. Then, $\textsc{co-NP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
\end{proof}

### 2015-05-12 tex/th3.tex
12 may

\begin{theo}
$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Since $\#\Sigma_1$ is monotonic under addition of elements, $f(\B) \geq 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}

>>>>

\begin{theo}
$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.

We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\Truc$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}. Therefore, $f(\B) \geq f(\A) = 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}
